---
title: "Selector Functionality"
author: John Mernitz
subtitle: Methods and Process
date: today
date-format: long
format:
  html:
    dark-mode: true
    theme: 
      light: minty
      dark: minty
    code-fold: true
    toc: true
---

# Underlying Principles

The shapefiles within this project are from the Town of Amherst's GIS division, and from the New York Natural Heritage Program(NYNHP). From a combination of the hydrographic and wetland layers of Amherst, a buffer layer was composed to encompass a quarter mile around those features. Similarly, a buffer around the NYNHP was created to provide an extended range wherein various plants of that biome would be recommended in priority.

# Database Organization

The plants data in this project is currently a product of David Werier and the New York Flora Association, who allowed permission of its use in this project. Plants therein are those documented by NYFA in their Atlas project as native to Erie County, albeit acknowledging that their list is not complete. The list of plants has the fields of scientific name, common name,family, genus, species, infraspecies, Coefficient of Conservatism, State Status, National Wetland Plant List code for Northcentral and Northeast, State Rank, Global Rank, Habitat, Growth Habit, and Duration. 

## Data Management

The data is divided into the above categories, but also must be split due to New York laws about threatened, endangered and rare species. It is forbidden to cause any bother or harm to these species, and so their data must be partitioned. 
```{r}
library(tidyverse)
plant_data <- read.csv("data/Copy_NYFA_Erie_Native_pared.csv")
#Downloading data
```

This first section divides the data based on threat status. Extirpated plants should be removed as well.

```{r}
plant_data <- plant_data %>%
  filter(plant_data[,14] != "SX")
         
common_plants <- plant_data %>%
  filter(!(plant_data[,12] %in% c("Endangered-State", "Threatened-State", "Rare-State")))
#first filter for non-threatened data

threatened_plants <- plant_data %>%
  filter(plant_data[,12] %in% c("Endangered-State", "Threatened-State", "Rare-State"))
#second filter for threatened plant species
```

Values should be assigned to the wetland codes of the plants, to better qualify their need for water.

Another few splits should be added here to partition plants which require more wet and more dry conditions.

## Calling Data

Layer files from the project are transformed into geojson for call expediency and ease of display.

## Display



# Site Aesthetics and Utility



# Linked Paper



# Citations



