---
title: "Selector Functionality"
author: John Mernitz
subtitle: Methods and Process
format:
  html:
    dark-mode: true
    theme: 
      light: minty
      dark: minty
    code-fold: true
    toc: true
---

# Underlying Principles

The shapefiles within this project are from the Town of Amherst's GIS division, and from the New York Natural Heritage Program(NYNHP). From a combination of the hydrographic and wetland layers of Amherst, a buffer layer was composed to encompass a quarter mile around those features. Similarly, a buffer around the NYNHP was created to provide an extended range wherein various plants of that biome would be recommended in priority.

# Database Organization

The plants data in this project is currently a product of David Werier and the New York Flora Association, who allowed permission of its use in this project. Plants therein are those documented by NYFA in their Atlas project as native to Erie County, albeit acknowledging that their list is not complete. The list of plants has the fields of scientific name, common name,family, genus, species, infraspecies, Coefficient of Conservatism, State Status, National Wetland Plant List code for Northcentral and Northeast, State Rank, Global Rank, Habitat, Growth Habit, and Duration. 

## Data Management

The data is divided into the above categories, but also must be split due to New York laws about threatened, endangered and rare species. It is forbidden to cause any bother or harm to these species, and so their data must be partitioned. 
```{r, message=F, warning=F, results=F}
library(tidyverse)
plant_data <- read.csv("data/Copy_NYFA_Erie_Native_pared.csv")
#Downloading data
```

This first section divides the data based on threat status. Extirpated plants should be removed as well.

```{r, message=F, warning=F, results=F}
plant_data <- plant_data %>%
  filter(plant_data[,14] != "SX")
         
common_plants <- plant_data %>%
  filter(!(plant_data[,12] %in% c("Endangered-State", "Threatened-State", "Rare-State")))
#first filter for non-threatened data

threatened_plants <- plant_data %>%
  filter(plant_data[,12] %in% c("Endangered-State", "Threatened-State", "Rare-State"))
#second filter for threatened plant species
```

Now we have a list of common plants we can export to csv.

```{r, message=F, warning=F, results=F}
write.csv(common_plants,"data/common_plants.csv", row.names = FALSE)
```

We also need to alter the new csv to make filtering possible.

```{r}
native_plants <- read.csv("data/common_plants.csv")

native2 <- native_plants[,c("Scientific_Name", "Common_Name", "CoC", "nwpl_ncne", "Growth_Habit", "Duration")]

#Fewer columns, edit the col names so it's more clear what each refers to

native2 <- native2 %>%
  rename(
    `Scientific Name` = Scientific_Name,
    Common = Common_Name,
    Difficulty = CoC,
    `Wetland Status` = nwpl_ncne,
    `Growth Habit` = Growth_Habit,
    Duration = Duration
  )

# Normalize and Un-nest the Growth Habit data 
native2_cleaned <- native2 %>%
  # Split cells with multiple values into separate rows
  separate_rows(`Growth Habit`, sep = ",|;\\s*") %>% 
  # Clean up whitespace/case again after splitting
  mutate(
    `Growth Habit` = str_trim(`Growth Habit`),
    `Growth Habit` = str_to_title(`Growth Habit`)
  ) %>%
  # Remove rows that ended up blank after the split/trim
  filter(!is.na(`Growth Habit`), `Growth Habit` != "")
  
write_csv(native2_cleaned, "D:/UB/Masters Project/MastersProjQuart/data/native2_cleaned.csv")

```
We can call the cleaned csv on the other page.

## Calling Data

Layer files from the project are transformed into geojson for call expediency and ease of display.
```{r}
ditches <- read_sf("https://drive.usercontent.google.com/u/0/uc?id=1SLCufnLDpR4bEJeUB1usJ2OFzBlWXtEY&export=download")
town_bound <- read_sf("https://drive.usercontent.google.com/u/0/uc?id=1WScoYQWLa9_jaobUhm-0xcf1MtTaqoE6&export=download")
state_wetld <- read_sf("https://drive.usercontent.google.com/u/0/uc?id=1uS0oYTaWLGhwGyYHnhkBxn8RZRlwNFkk&export=download")
#Wetland makes up 20.2228% of area of Amherst NY
soils_C <- read_sf("https://drive.usercontent.google.com/u/0/uc?id=1zItxfisuLYZP6ZhnyUzMGH80UShGLAUQ&export=download")
nysnh_cl <- read_sf("https://drive.usercontent.google.com/u/0/uc?id=1Mjfgt23T8KmxcY5OzBJ0DZ4dgW2jnoHd&export=download")
hydro_cl <- read_sf("https://drive.usercontent.google.com/u/0/uc?id=13OIxXMGIlrMYLmIUJuhU8ElM9Dpkoz11&export=download")
wetldsum <- rast("data/WtldSumVal1_8.tif")
#wetldunion <- read_sf("https://drive.usercontent.google.com/u/3/uc?id=1yNnPViAN8-cjZ38yrhCgsEoaqXd_Un9z&export=download")
#Download.file()
#Save to tempdata folder

town_bound <- st_transform(town_bound, crs = 4326)
state_wetld <- st_transform(state_wetld, crs = 4326)
soils_C <- st_transform(soils_C, crs = 4326)
ditches <- st_transform(ditches, crs = 4326)
nysnh_cl <- st_transform(nysnh_cl, crs = 4326)
hydro_cl <- st_transform(hydro_cl, crs = 4326)
wetldsum <- project(wetldsum, "EPSG:4326")

soil_names <- (unique(soils_C$SOIL_NAME))
wetldsumpal <- colorNumeric(palette = "viridis", domain = values(wetldsum), na.color = "transparent")
soils_pal <- colorFactor(palette = "turbo", domain = soil_names, na.color = "transparent")
```

## Display

The following is the first attempt at a map of wetland proximity.

```{r, fig.width=6, fig.heigt=3, fig.cap="Map of Wetland Files"}
m <- leaflet(town_bound) %>%
  addTiles() %>% # Add a default base layer
  addPolygons(data = town_bound, group = "Amherst Boundary", stroke = TRUE, color = "black", fill = FALSE, weight = 3) %>%
  addPolygons(data = soils_C, group = "Soil Types", color = "gray", weight = 1, fillOpacity = 0.5, fillColor = ~soils_pal(SOIL_NAME), label = ~SOIL_NAME,
    highlightOptions = highlightOptions(weight = 3, color = "white", bringToFront = FALSE)) %>%
  addPolygons(data = nysnh_cl, group = "Significant Wetland", color = "gray", weight = 1, fillOpacity = 0.5, fillColor = "yellow") %>%
  addPolygons(data = hydro_cl, group = "Hydrology", color = "gray", weight = 1, fillOpacity = 0.5, fillColor = "blue") %>%
#  addHomeButton(
#    ext = town_bound,
#    group = "Home",
#    position = "bottomright"
#  ) %>%
  addRasterImage(wetldsum, group = "Wetland Proximity Values", colors = wetldsumpal, opacity = 0.8) %>%
  #addImageQuery(wetldsum, project = TRUE, layerId = "Wetldsum", position = "top") %>%
  addLegend(pal = wetldsumpal, values = values(wetldsum), title = "Wetland Proximity values", opacity = 1) %>%
#  addScaleBar(data=town_bound, position = c("topright", "bottomright", "bottomleft", "topleft"),
#  options = scaleBarOptions()
#)

#scaleBarOptions(
#  maxWidth = 100,
#  metric = TRUE,
#  imperial = TRUE,
#  updateWhenIdle = TRUE
#)
  #addLegend(pal = soils_pal, values = soils_C$SOIL_NAME, title = "Soil Types", position = "bottomright", opacity = 1)
  #addImageQuery(wetldsum, type = "mousemove", digits = 2) %>%
  addLayersControl(
    baseGroups = c("OSM (default)"),
    overlayGroups = c("Amherst Boundary", "State Wetland", "Soil Types", "Significant Wetland", "Hydrology", "Ditches", "Wetland Proximity Values"),
    options = layersControlOptions(collapsed = FALSE)
  )

m

```


# Site Aesthetics and Utility



# Linked Paper



# Citations



