---
title: "Geographies of Native Plant Restoration"
author: John Mernitz
subtitle: Planting Selection Tool
date: today
date-format: long
format:
  html:
    dark-mode: true
    theme: 
      light: minty
      dark: minty
    code-fold: true
    toc: true
---

# Introduction

Welcome to the Plant Selection tool! If you would like to know more about this tool and its functionality, please check the about and code steps on the navbar above!

Again, this plant selector is not fully tested, plants should be carefully considered if they are appropriate for one's needs. This page seeks to diversify thoughts and choices about native plants in gardens; do not blindly choose one. Please be aware some of these plants have natural defenses like thorns, irritating sap, or inedible fruits!

Suggestions are organized by hydrologic and wetland proximity, annual or perennial, and by growth type.

# Leaflet Map

```{r, message=F, warning=F}
library(tidyverse)
library(leaflet)
library(kableExtra)
library(htmlwidgets)
library(widgetframe)
knitr::opts_chunk$set(widgetframe_widgets_dir = 'widgets' ) 
knitr::opts_chunk$set(cache=TRUE)  # cache the results for quick compiling
```

## Download and clean all required data

```{r}
n=20
data=data.frame(x=runif(n,-180,180),
                y=runif(n,-60,60),
                size = runif(n, 5, 20),
                category = factor(
                  sample(letters[1:5], n, replace = TRUE)
                  ),
                value = rnorm(n))
```

```{r, results='asis'}
data %>% 
  slice(1:10) %>% #show only 1:n rows
  kable(digits=2,align="c")%>% #make table and round to two digits
  kable_styling(bootstrap_options = 
                  c("striped", "hover", "condensed", "responsive")) #apply other formatting
```

```{r, fig.width=6, fig.height=3, fig.cap="Map of completely random data"}
m <- leaflet(data) %>% 
  addTiles() %>% 
  addCircleMarkers(~x, ~y, radius = ~size,color = ~as.factor(category)) %>% 
  addPopups(~x[2], ~y[2], "Random popup")
m  # a map with the default OSM tile layer
```
