---
subtitle: "**Environment Locator and Plant Selector**"
format:
  html:
    dark-mode: true
    theme: 
      light: minty
      dark: minty
    code-fold: true
    toc: true
    page-layout: full
editor_options: 
  chunk_output_type: console
---

```{=html}
<style>
.cs-filter-grid .crosstalk-options-group {
  display: grid;
  grid-template-columns: repeat(auto-fill, minmax(150px, 1fr));   
  grid-gap: 10px;
  margin-top: 10px;
}
.cs-filter-grid .checkbox {
  margin: 0;               
  display: flex;
  align-items: flex-start;
  break-inside: avoid;
}
.cs-filter-grid .checkbox input {
  margin-right: 8px;
  margin-top: 4px;
}
.column-sm-3 {
  padding-right: 25px !important; 
}
.control-label {
  font-weight: bold;
  margin-bottom: 5px;
}
@media (max-width: 768px) {
  /* Fix: Ensure the class name matches your R code */
  .cs-filter-grid .crosstalk-options-group {
    grid-template-columns: 1fr;
  }
  .column-sm-3 {
    margin-bottom: 30px;
    padding-right: 15px !important;
  }
}
</style>
```

```{r echo=FALSE, message=FALSE, warning=FALSE, results=F}
#install.packages("reactable")
#install.packages("sf")
#install.packages("terra")
#install.packages("leaflet")
library(tidyverse)
library(leaflet)
library(kableExtra)
library(htmlwidgets)
library(htmltools)
library(widgetframe)
library(dplyr)
library(reactable)
library(crosstalk)
library(sf)
library(mapview)
library(leafem)
library(terra)
knitr::opts_chunk$set(widgetframe_widgets_dir = 'widgets' ) 
knitr::opts_chunk$set(cache=FALSE)  # cached results off
```

**Use this page to help select native plant species for your local environment!**

**How to Use This Tool:**

i)  **Navigate to your property** or neighborhood in the Town of Amherst. Look at the Soil and Wetland Value layer to get a sense of your site's conditions.

ii) **Match your conditions:** Use the displayed Wetland Value and personal knowledge of your site to choose the values you would like to explore in the plant list

iii) **Browse Native Plants:** Use the filters to explore which native plants would work best for your conditions

**Step 1:** On the map below, you can find your location and corresponding soil type/wetland value. Layers can be toggled in the top-right corner of the map!

```{r echo=FALSE}
town_bound <- read_sf("https://drive.usercontent.google.com/u/0/uc?id=1WScoYQWLa9_jaobUhm-0xcf1MtTaqoE6&export=download")

nysnh_cl <- read_sf("https://drive.usercontent.google.com/u/0/uc?id=1Mjfgt23T8KmxcY5OzBJ0DZ4dgW2jnoHd&export=download")

wetldsum <- rast("data/WtldSumVal1_8.tif")
wetldunion <- read_sf("https://drive.usercontent.google.com/u/3/uc?id=1yNnPViAN8-cjZ38yrhCgsEoaqXd_Un9z&export=download")

town_bound <- st_transform(town_bound, crs = 4326)
nysnh_cl <- st_transform(nysnh_cl, crs = 4326)
wetldunion <- st_transform(wetldunion, crs = 4326)
wetldsum <- project(wetldsum, "EPSG:4326")

soil_names <- (unique(wetldunion$SOIL_NAME))

bins1 <- 0:10

wetldsumpal <- colorNumeric(palette = "viridis", domain = values(wetldsum), na.color = "transparent")

wetldpallette <- colorBin(palette = "inferno", domain = wetldunion$Wetld_Wgt,  bins = bins1, na.color = "transparent")
wetldlabel<-  paste("This section has a <b>Wetland Value</b> of <b>", wetldunion$Wetld_Wgt, "<br/>", "Soil</b>", wetldunion$HYDRICSTAT, "<br/>", wetldunion$SOIL_NAME)
#soils_pal <- colorFactor(palette = "turbo", domain = soil_names, na.color = "transparent")
```

```{r, echo=FALSE, fig.width=6, fig.heigt=3, fig.cap="Map of Wetland Files"}
m <- leaflet(town_bound) %>%
  addTiles() %>% # Add a default base layer
  addPolygons(data = town_bound, group = "Amherst Boundary", stroke = TRUE, color = "black", fill = FALSE, weight = 3) %>%
  addPolygons(data = wetldunion, group = "Soil and Wetland Value", color = "gray", weight = 1, fillOpacity = 0.5, fillColor = ~wetldpallette(Wetld_Wgt), label = lapply(wetldlabel, htmltools::HTML),highlightOptions = highlightOptions(weight = 3, color = "white", bringToFront = TRUE)) %>%
  addPolygons(data = nysnh_cl, group = "Significant Wetland", color = "gray", weight = 1, fillOpacity = 0.5, fillColor = "yellow") %>%
#  addHomeButton(
#    ext = town_bound,
#    group = "Home",
#    position = "bottomright"
#  ) %>%
  addRasterImage(wetldsum, group = "Wetland Proximity", colors = wetldsumpal, opacity = 0.8) %>%
  addLegend(position = "bottomright", group = "Value Legend", pal = wetldpallette, values = wetldunion[[4]], title = "Wetland Value (0-10)", opacity = 1) %>%
  addScaleBar(position = c("bottomleft"), options = scaleBarOptions(maxWidth = 100, imperial = TRUE, updateWhenIdle = FALSE)) %>%
  addLayersControl(
    baseGroups = c("OSM (default)"),
    overlayGroups = c("Amherst Boundary", "Significant Wetland", "Soil and Wetland Value", "Wetland Proximity", "Value Legend"),
    options = layersControlOptions(collapsed = TRUE)
  )%>%
hideGroup("Wetland Proximity")
m

```

```{r echo=FALSE}
native2_cleaned <- read.csv("data/native2_cleaned.csv")
# Create SharedData object from the CLEANED data

# beware error due to unclear growth habit names
native3 <- SharedData$new(native2_cleaned)

wetland_labels <- c(
  "FAC"  = "Facultative (5-6)", 
  "FACU" = "Facultative Upland (3-4)", 
  "FACW" = "Facultative Wetland (7-8)", 
  "OBL"  = "Obligate Wetland (9-10)", 
  "UPL"  = "Upland (0-2)"
)
```

**Step 2: Take note of the Wetland Codes below!**

| **Wetland Values** | **Wetland Status Indicator** | Description |
|:--:|:--:|----|
| 0-2 | Upland (UPL) | Almost never present in wetlands |
| 3-4 | Facultative Upland (FACU) | Usually non-wetland, but may occur in wetlands |
| 5-6 | Facultative (FAC) | Present both in and out of wetlands |
| 7-8 | Facultative Wetland (FACW) | Usually found present in wetlands, but may survive non-wetland conditions |
| 9-10 | Obligate (OBL) | Almost always found present in wetlands |

: *Wetland Plant Classes as defined in National Wetland Plant List, values as calculated for this project*

**Step 3: Use the Wetland Values to estimate your location's Wetland Status. (e.g. Value of 0-2, choose UPL)**

**Plant Selection Reminder:** Plants should be carefully considered if they are appropriate for one's needs. This page seeks to diversify thoughts and choices about native plants in gardens; do not blindly choose one. Please be aware some of these plants have natural defenses like thorns, irritating sap, or inedible fruits!

Suggestions can be filtered by scientific name, common name, wetland status, growth type, and duration.

*Read more about why soil type and the above index matter on the **"[How it Works](https://johnmernitz.github.io/Native_Wetland_Plant_Selector/fnality.html)"** page!*

```{r echo=FALSE, interactive-table}
#| column: screen
bscols(
  widths = c(3,9),
  device = "sm",
  list(
    filter_select("Wtd_St", "Wetland Status", native3, group = ~wetland_labels[Wetland.Status]),
    tags$div(class = "cs-filter-grid",
      filter_checkbox("Gth_Hbt", "Growth Habit", native3, ~Growth.Habit)
      ),
    filter_slider("CoC", "Difficulty", native3, ~Difficulty, width = "100%"),
    filter_select("Drt", "Duration", native3, ~Duration)
  ),
  reactable(
    native3,
    outlined = TRUE,
    borderless = FALSE,
    striped = TRUE,
    searchable = TRUE,
    columns = list(
      Scientific.Name = colDef(name = "Scientific Name"),
    Common = colDef(name = "Common Name"),
    Difficulty = colDef(name = "Difficulty"),
    Wetland.Status = colDef(name = "Wetland Status"),
    Growth.Habit = colDef(name = "Growth Habit"),
    Duration = colDef(name = "Duration")
    ),
    minRows = 10,
    fullWidth = TRUE,
    defaultPageSize = 10,
    showPageSizeOptions = TRUE,
    pageSizeOptions = c(10, 25, 50, 100)
)
)

```

*Information regarding current data sources and last acquisition can be found on the **"[How it Works](https://johnmernitz.github.io/Native_Wetland_Plant_Selector/fnality.html)"** page as well. Project last updated as of May 2026.*
